---
title: "Demo 1: Analog Ensemble Basics"
author: "Weiming Hu"
date: "October 29, 2018"
output: html_document
# output: 
#   md_document:
#     variant: markdown_github
---

```{r setup, include=F}
library(RAnEn)
library(RAnEnExtra)
library(maps)

stopifnot(packageVersion('RAnEn')>="3.5.0")
```

## Introduction

This article walks you through the basic usage of the `RAnEn` library. This exercise uses short-term surface temperature forecasts as an example. Recommend using [binder](https://mybinder.org/v2/gh/Weiming-Hu/AnalogsEnsemble/master?urlpath=rstudio) and the corresponding `.Rmd` file will guide you through the script line by line.

You will learn how to use these functions:

- `generateConfiguration`
- `generateAnalogs`
- `verify*` functions

## Data Description

First, let's load the prepared NAM model forecasts and analysis. You can find the data in this direct folder if you want to run this script locally.

```{r}
file.name <- 'data-NAM-StateCollege.RData'
# file.name <- 'data-NAM-Denver.RData'
# file.name <- 'data-METAR-CONUS.Rdata'

if (!file.exists(file.name)) {
  cat('Downloading from the data server which might be slow ...\n')
  download.file(url = paste('https://prosecco.geog.psu.edu/', file.name, sep = ''),
                destfile = file.name)
}
load(file.name)
rm(file.name)

print(ls())
```

The following parameters are included and provided in the dataset.

```{r}
print(forecasts$ParameterNames)
```

Foreacsts and analysis have the following time span. Because the original time values record the number of seconds since an original time, the function `as.POSIXct` is used to convert them to readable date-time class objects. Alternatively, `RAnEn::toDateTime` can complete the same action. It is a convenient wrapper function of `as.POSIXct`.

```{r}
print(as.POSIXct(range(forecasts$Times), origin = '1970-01-01', tz = 'UTC'))
print(as.POSIXct(range(observations$Times), origin = '1970-01-01', tz = 'UTC'))
```

Coordinates in lat/lon are provided for each grid point. We can plot them on maps. **Note that coordinates start from the bottom left and proceed in a row-wise order**. There are 11 rows and 11 columns of grid points in the data sets.

```{r}
map('county', regions = 'pennsylvania', col = 'grey')
points(forecasts$Xs - 360, forecasts$Ys, pch = 16, cex = .3, col = 'red')
```

Finally, let's look at the dimensions of forecasts and analysis.

```{r}
print(dim(forecasts$Data))
print(dim(observations$Data))
```

Forecasts have 4 dimensions which include parameters, stations/grid points, days, and Forecast Lead Times (FLTs). Analysis have 3 dimensions which include parameters, stations/grid points, and time.

## Generate Temperature Forecasts

Now that we have a general idea of the data, we can generate temperature forecasts. The steps to generate AnEn forecasts are as follow:

- Initialize a configuration
- Set up options
- Generate AnEn

First, we set up the configuration. We generate an AnEn for one day using about one year of historical data for all the grid points available. For a full list of avaiable options, please check the document for [RAnEn::generateConfiguration](https://weiming-hu.github.io/AnalogsEnsemble/R/reference/generateConfiguration.html).

```{r}
# We then generate a configuration
config <- new(Config)

# Set up the start and end indices for test times
test.start <- 366
test.end <- 372

# Set up the start and end indices for search times
search.start <- 1
search.end <- 365

# # Set the test times to generate AnEn for
test.times <- forecasts$Times[test.start:test.end]
# 
# # Set the search times to generate AnEn from
search.times <- forecasts$Times[search.start:search.end]

# Set up the number of members to keep in the analog ensemble
# Empirically, the number of members is the square root of number of search times.
config$num_analogs <- sqrt(search.end - search.start + 1)

# Set up the variable that we want to generate AnEn for.
# This is the index of parameter in observation parameter names.
#
config$observation_id <- 8

# Set up weights for each parameters
config$weights <- c(1, 1, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 1)

```

Now we can generate an AnEn. This could take a while. On a Mac Air, it takes about 2 minutes.

```{r}
AnEn <- generateAnalogs(forecasts, observations, test.times, search.times, config)
```

The result variable `AnEn` is a list of members. Let's check what the members are.

```{r}
print(AnEn)
```

## Verification

Finally, let's extract the AnEn forecast values and the corresponding observations.

```{r}
# Collect analog ensemble values
anen <- AnEn$analogs

# Collect observations for the test period and the forecasted weather variable
obs <- alignObservations(observations$Data, observations$Times,
                         test.times, forecasts$FLTs,
                         silent = T, show.progress = F)
obs <- obs[config$observation_id, , , ]
```

Then we can generate verification statistics, and see how AnEn is performing.

``` {r}
# Generate verification metrics
ret.MAE <- verifyMAE(anen, obs)
ret.RMSE <- verifyRMSE(anen, obs)
ret.Bias <- verifyBias(anen, obs)
ret.RH <- verifyRankHist(anen, obs)

# Let's make some figures
par(mfrow = c(4, 1), mar = c(3, 4.5, 1, 1))
plot(forecasts$FLTs/3600, ret.MAE$flt, type = 'b', pch = 1, cex = 0.5,
     xlab = '', ylab = 'MAE')
plot(forecasts$FLTs/3600, ret.RMSE$flt, type = 'b', pch = 1, cex = 0.5,
     xlab = '', ylab = 'RMSE')
plot(forecasts$FLTs/3600, ret.Bias$flt, type = 'b', pch = 1, cex = 0.5,
     xlab = 'Lead Times (h)', ylab = 'Bias')
barplot(ret.RH$rank, ylab = 'Rank Frequency')
```

