---
title: "Demo 3: Analog Ensemble with Operational Search"
author: "Weiming Hu"
date: "February 12, 2019"
output: html_document
# output:
#   md_document:
#     variant: markdown_github
---

```{r setup, include=FALSE}
library(RAnEn)
library(RAnEnExtra)
library(maps)

stopifnot(packageVersion('RAnEn')>="4.0.2")
```

## Introduction

Prediction accuracy of the Analog Ensemble depends on the quality of analogs. Presumably, better analogs will generate better predictions. In an operational model, it is likely that the historical forecasts in the near past are the most similar to the current forecast. Therefore, in operational mode, as each day passes, it is added to the historical repository.

This article shows an example of how to use `RAnEn` with an operational search. It is strongly suggested to go over the [demo 1](https://weiming-hu.github.io/AnalogsEnsemble/2018/11/04/demo-1-RAnEn-basics.html) prior to this tutorial.

## Data Description

Same data are used from the [demo 1](https://weiming-hu.github.io/AnalogsEnsemble/2018/11/04/demo-1-RAnEn-basics.html).

```{r}
file.name <- 'data-NAM-StateCollege.RData'
# file.name <- 'data-NAM-Denver.RData'

if (!file.exists(file.name)) {
  cat('Downloading from the data server which might be slow ...\n')
  download.file(url = paste('https://prosecco.geog.psu.edu/', file.name, sep = ''),
                destfile = file.name)
}

load(file.name)
rm(file.name)
```

## Generate Temperature Forecasts

When generating forecasts with an operational search, `search_times_compare` is not necessary because for each test time in the `test_times_compare`, the search times will be automatically calculated and they include the historical times until the day before the test time.

```{r}
# Set test and search times
test.start <- 366
test.end <- 372
search.start <- 1
search.end <- test.start - 1

test.times <- forecasts$Times[test.start:test.end]
search.times <- forecasts$Times[search.start:search.end]

config <- new(Config)
config$num_analogs <- 19

# Setting them to large numbers to allow NA
config$max_flt_nan <- 10000
config$max_par_nan <- 10000

# Only save the similarity time indices
config$save_similarity_time_index <- T
config$save_analogs <- F

# Save all similarity
config$num_similarity <- length(search.start:search.end)

# Don't forget to trigger the operational search
config$operation <- T
```

We then generate analogs. This can take a while. On a Mac Air, it takes about 4 minutes.

```{r}
AnEn.opr <- generateAnalogs(forecasts, observations, test.times, search.times, config)
```

## Check Search Space for Operational Mode

If we take a close look at the analog members of operational search. The search times of the operational search for test times are different, actually increasing. This is indeed the results from the operational search.

```{r}
# Convert times to POSIX type so that it is easier to read
test.times <- as.POSIXct(test.times, origin = '1970-01-01', tz = 'UTC')
fcst.times <- as.POSIXct(forecasts$Times, origin = '1970-01-01', tz = 'UTC')

i.test.day <- 1
cat('The search time range for', format(test.times[i.test.day], format = '%Y/%m/%d'), ':', 
    format(range(fcst.times[AnEn.opr$similarity_time_index[1, i.test.day, 1, ]], na.rm = T), format = '%Y/%m/%d'),
    ' Total count:', length(which(!is.na(AnEn.opr$similarity_time_index[1, i.test.day, 1, ]))), '\n')

i.test.day <- 2
cat('The search time range for', format(test.times[i.test.day], format = '%Y/%m/%d'), ':', 
    format(range(fcst.times[AnEn.opr$similarity_time_index[1, i.test.day, 1, ]], na.rm = T), format = '%Y/%m/%d'),
    ' Total count:', length(which(!is.na(AnEn.opr$similarity_time_index[1, i.test.day, 1, ]))), '\n')
```
